### 2020-10-21 manuscript.Rmd
First draft of introduction

---
documentclass: jss
classoption: nojss
citation_package: natbib
bibliography: references.bib
author:
  - name: Devin Incerti
    affiliation: Genentech
    # use this syntax to add text on several lines
    address: |
      | South San Francisco
    email: \email{incerti.devin@gene.com}
  - name: Jeroen Jansen
    affiliation: |
      | University of California, San Francisco
    address: |
      | San Francisco
    email: \email{jeroen.jansen@ucsf.edu}
title:
  plain:     "hesim: Health Economic Simulation Modeling and Decision Analysis"
  formatted: "\\pkg{hesim}: Health Economic Simulation Modeling and Decision Analysis"
abstract: >
  The abstract of the article.
keywords:
  formatted: [health economic evaluation, cost-effectiveness analysis, simulation, modeling, "\\proglang{R}"]
  plain: [health economic evaluation, cost-effectiveness analysis, simulation, modeling, R]
preamble: >
  \usepackage{amsmath}
output: rticles::jss_article
---

@@

# Introduction

>>>>

---
documentclass: jss
classoption: nojss
citation_package: natbib
bibliography: references.bib
author:
  - name: Devin Incerti
    affiliation: Genentech
    # use this syntax to add text on several lines
    address: |
      | South San Francisco
    email: \email{incerti.devin@gene.com}
  - name: Jeroen Jansen
    affiliation: |
      | University of California, San Francisco
    address: |
      | San Francisco
    email: \email{jeroen.jansen@ucsf.edu}
title:
  plain:     "hesim: Health Economic Simulation Modeling and Decision Analysis"
  formatted: "\\pkg{hesim}: Health Economic Simulation Modeling and Decision Analysis"
abstract: >
  To do.
keywords:
  formatted: [health economic evaluation, cost-effectiveness analysis, simulation, modeling, "\\proglang{R}"]
  plain: [health economic evaluation, cost-effectiveness analysis, simulation, modeling, R]
preamble: >
  \usepackage{amsmath}
output: rticles::jss_article
---

@@

# Introduction
Health technology assessment (HTA) is a systematic approach for comparing competing health technologies to inform the efficient use of health care resources. Publicly funded health systems such as those in Australia, Canada, and the United Kingdom, among others, use HTA to help maximize health gains for the population given a fixed budget. HTA is also commonly used by private payers to guide coverage decisions and the adoption of new technologies \citep{trosman2011health}. Such assessments usually rely heavily on cost-effectiveness analysis (CEA) so that decisions can be made on the basis of a formal evaluation of costs and effects \citep{dakin2015influence}. CEA is made feasible by development of health economic models that simulate costs and effects over relevant time horizons using the totality of available evidence.

The most commonly used economic models are Markov state transition models (STMs) that simulate transitions between mutually exclusive health states. When the Markov assumption holds so that transition probabilities are either constant over time or depend only on model time, a cohort-level model can be used \citep{briggs1998introduction}. If the Markov assumption is relaxed---e.g., with a semi-Markov model so that transition probabilities depend on time in an intermediate health state---then individual-level models are required in most cases \citep{brennan2006taxonomy, fiocco2008reduced}. Individual-level models afford considerably more flexibility and allow patient history to be tracked over time.

Decisions informed by economic models and CEA are subject to uncertainty. One source of decision uncertainty stems from uncertainty in the underlying model parameters. Parameter uncertainty is typically quantified using probabilistic sensitivity analysis (PSA), which involves randomly sampling the model parameters from suitable probability distribution and simulating the model for each sampled parameter set \citep{claxton2005probabilistic}. When combined with individual-level simulation, PSA can take an appreciable amount of time to run \citep{ohagan2007monte}.

Estimation of model parameters should ideally be performed using statistical models that are aligned with the structure of the economic model. For instance, when patient-level data is available, a multi-state model can be used to parameterize all possible transitions in a STM while accounting for censoring and competing risks \citep{williams2017cost}. Similarly, in oncology, partitioned survival models (PSMs) can be parameterized from estimates of progression-free survival (PFS) and overall survival (OS). In other cases, parameters might be combined from disparate sources, such as within a single Bayesian model \citep{baio2012bayesian}. When the clinical evidence base is not limited to a single study, a formal evidence synthesis, such as a network-meta analysis (NMA), might even be performed \citep{dias2018network}.

Despite their computational demands and foundations in statistics, health economic models have historically been developed with specialized commercial software (e.g., \proglang{TreeAge}) or more commonly with a spreadsheet (almost always Microsoft \proglang{Excel}). The limitations of such software relative to programming languages like \proglang{R} have been increasingly emphasized in the literature \citep{baio2017simple, incerti2019r, jalal2017overview}. It is therefore no surprise that a number of related \proglang{R} packages have recently been developed, such as \pkg{BCEA} \citep{baio2017bayesian}, \pkg{SAVI} \citep{strong2014estimating}, \pkg{survHE} \citep{baio2020survhe}, and \pkg{heemod} \citep{filipovic2017markov}. Still, of the available packages, only \pkg{heemod} provides a general purpose framework for developing simulation models and it is limited to cohort Markov models.

`hesim` is an \proglang{R} package that advances the functionality and performance of the existing software. Multiple model types are supported including cohort discrete time state transition models (DTSTMs), N-state PSMs, and individual-level continuous time state transition models (CTSTMs), encompassing both Markov (time-homogeneous and time-inhomogeneous) and semi-Markov processes. To maximize flexibility and faciliate integration of the statistical methods and economic model, parameters can be estimated either by fitting a model in \proglang{R} or by inputting parameters obtained from external sources. So that individual-level simulation and PSA can be run quickly, \pkg{Rcpp} and \pkg{data.table} are heavily utilized. After simulating costs and quality-adjusted life-years (QALYs) from a PSA, decision analysis can be performed within a cost-effectiveness framework.
